# Methodology

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
library(here)
library(osmdata)
library(sf)
library(tidyverse)
library(ggmap)
library(leaflet)
library(tidycensus)
library(modelsummary)
library(qwraps2)
library(knitrProgressBar)
library(lwgeom)

basetiles <- providers$CartoDB.Positron
this_crs <- 2227 # EPSG:2227 - NAD83 / California zone 3 (ftUS)
options(tigris_use_cache = TRUE)
```
This data analysis of this project involves two main parts. The first part is the data collection of grocery stores and their locations in Utah county. The first code to create was a code that pulled the grocery store data from open street map data source.
```{r}
bb <- getbb("Utah County, Utah", format_out = "polygon")
supermarket_osm <- opq(bb) %>% #specify bounding query
  add_osm_feature(key = "shop", value = "supermarket") %>% #specify which kinds of data we want
  osmdata_sf() %>% # get a list of sf data frames for these tags
  trim_osmdata(bb, exclude = TRUE)
```

The next step was to filter out the polygons from the points data in the set and then transform the polygons into points so that all the data could be the same type. We then organized the data according to their osm_id.  
```{r}
polygons <- supermarket_osm$osm_polygons %>% st_transform(this_crs)

points_on_polygon <- supermarket_osm$osm_points %>%
  st_transform(this_crs)

supermarket <- rbind(
  #convert polygons to centroids
  polygons %>% st_centroid() %>%
    select(osm_id, name),
  # get points
  supermarket_osm$osm_points %>%
    st_transform(this_crs) %>%
    filter(!(osm_id %in%
               points_on_polygon$osm_id)) %>%
    select(osm_id, name = brand)
)
```

The last part of this data analysis set was to present this data in a visual manner, which was accomplished using a leaflet tool, as shown below. 
```{r}

leaflet(supermarket %>% st_transform(4326)) %>%
  addProviderTiles(basetiles) %>%
  addCircleMarkers(color = "red")
```

The second part of data analysis involves collecting data about block groups in the Utah County. This data comes from the Census Bureau and the first step is to draw the data and organize it according to race and income. 
```{r}

variables <- c(
  "population" = "B25008_001", # TOTAL POPULATION IN OCCUPIED HOUSING UNITS BY TENURE
  "housing_units" = "B25001_001", # HOUSING UNITS
  "households" = "B19001_001", #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  #Estimate!!Total!!Female!!Worked in the past 12 months!!Usually worked 35 or more hours per week
  # RACE
  "black" = "B02001_003",
  "asian" = "B02001_005",
  "pacific" = "B02001_006",
  "nativeam" = "B02001_004",
  "other" = "B02001_007",
  # HISPANIC OR LATINO ORIGIN BY SPECIFIC ORIGIN
  # The number of hispanic individuals needs to be drawn from a different table.
  # But this is only available at the tract level, where it appears to be roughly
  # collinear with the "some other race alone"
  "hispanic" = "B03001_003",
  #MEDIAN HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "income" = "B19013_001",
  #HOUSEHOLD INCOME IN THE PAST 12 MONTHS (IN 2017 INFLATION-ADJUSTED DOLLARS)
  "inc_0010" = "B19001_002",  "inc_1015" = "B19001_003", "inc_1520" = "B19001_004",
  "inc_2025" = "B19001_005", "inc_2530" = "B19001_006", "inc_3035" = "B19001_007",
  "inc_125"  = "B19001_015", "inc_150"  = "B19001_016", "inc_200"  = "B19001_017"
)

acs_bg <- get_acs(geography = "block group", variables = variables, year = 2018,
                  state = "UT", county = "049", geometry = TRUE)
acs <- acs_bg %>%
  select(-moe) %>%
  spread(variable, estimate) %>%
  mutate(area = as.numeric(st_area(geometry) * 1e-6)) %>%
  select(-hispanic) %>%
  # area is in m^2, change to km^2
  transmute(
    geoid = GEOID,
    group = 1,
    population, households, housing_units, 
    density = households / area,
    income,
    # many of the variables come in raw counts, but we want to consider
    # them as shares of a relevant denominator.
    lowincome    = 100 * (inc_0010 + inc_1015 + inc_1520 + inc_2530 +
                            inc_3035) / households,
    highincome   = 100 * (inc_125 + inc_150 + inc_200) / households,
    black        = 100 * black / population,
    asian        = 100 * asian / population,
    other        = 100 * (nativeam + pacific + other) / population,
    minority     = black + other
  )%>%
  filter(population > 0)

```
```{r}
leaflet(acs %>% st_transform(4326)) %>%
  addProviderTiles(basetiles) %>%
  addPolygons(data = acs, weight = 2, color = "red")
```
In order to combine both these data sets, we start by creating a half mile radius around each of the grocery stores. This half mile radius is what was decided on for an accessibility distance.

```{r}
supermarket_buffer <- st_buffer(supermarket, dist = 805) #this distance is in meters, but is approximately half a mile

sup_buff_int <- st_intersection(supermarket, supermarket_buffer)

bb_buff <- st_bbox(sup_buff_int)
```

We then intersect the larger radius grocery store data set with the block group data set in order to find the block groups that are within half a mile of a grocery store.
```{r}
sup_crs <- st_transform(supermarket_buffer, 4326) #make sure the crs is the same for both data sets
bg_crs <- st_transform(acs, 4326)
bg_int <- st_intersection(sup_crs, bg_crs)

```
After the two data sets were intersected we are left with a smaller block group data set that shows all block groups that were within the half mile radius of a grocery store. These block groups that are left are shown in a leaflet as well as a table that identifies the block groups and their different variables and features. 
```{r}
leaflet(bg_int %>% st_transform(4326)) %>%
  addProviderTiles(basetiles) %>%
  addPolygons(data = bg_int, weight = 2, color = "red")

```
A table showing the block group feature and identification is shown as well.
```{r}
bg_table <- bg_int %>% select(
    "Density: Households per square kilometer" = density,
    "Income: Median tract income" = income,
    "Low Income: Share of households making less than $35k" = lowincome,
    "High Income: Share of households making more than $125k" = highincome,
    "Black: Share of population who is Black" = black,
    "Asian: Share of population who is Asian" = asian,
    "Other: Share of population who belong to other minority groups" = other)

if(knitr::is_latex_output()){ 
  datasummary_skim(bg_table, title = "Block Group Summary Statistics",
    booktabs = TRUE, histogram = FALSE, output = "latex") %>%
    kableExtra::column_spec(1, width = "3cm")
} else {
  datasummary_skim( bg_table, title = "Block Group Summary Statistics")
}
```
